\section{Appendix}
\label{sec:appendix}

\subsection{Model Architecture Details}
\label{sec:appendix:arch}

gpt-oss-20b uses GptOssTopKRouter with a linear projection
$\mathbb{R}^{2880} \to \mathbb{R}^{32}$ and per-expert bias, followed
by top-4 selection with sigmoid normalization. Experts are implemented
as GptOssExperts with packed weight tensors: gate\_up\_proj of shape
$(32, 2880, 5760)$ and down\_proj of shape $(32, 2880, 2880)$, using
SwiGLU activation. These implementation details were discovered via
runtime inspection using \texttt{scripts/explore\_model.py}.

\subsection{Extraction MLP Architecture}
\label{sec:appendix:extraction}

Each extraction probe uses the following architecture:
\begin{verbatim}
Linear(2880, 512) -> SiLU -> Dropout(0.1)
Linear(512, 128)  -> SiLU -> Dropout(0.05)
Linear(128, 1)
\end{verbatim}
Total parameters per probe: $2880 \times 512 + 512 + 512 \times 128 +
128 + 128 \times 1 + 1 = 1{,}540{,}865$.
Training: AdamW optimizer, learning rate $10^{-3}$, weight decay 0.01,
cosine annealing schedule, 300 epochs, batch size 64, Huber loss on
log-scaled targets. 5-fold cross-validation with 3 seeds per fold;
best seed selected by validation $R^2$. Final ensemble: 5 probes
(one per fold), predictions aggregated via median.

\subsection{Router Signal Classifier}
\label{sec:appendix:router}

\begin{verbatim}
Linear(2880, 128) -> SiLU -> Dropout(0.1)
Linear(128, 32)   -> SiLU
Linear(32, 1)     -> Sigmoid
\end{verbatim}
Trained with BCE loss, 30 epochs, learning rate $10^{-3}$.
Achieves 100\% accuracy on held-out validation.

\subsection{Dataset Generation}
\label{sec:appendix:data}

Training and evaluation scenarios are generated procedurally with the
following distributions:
\begin{itemize}
\item \textbf{Property types}: 8 types sampled uniformly
(${\sim}$12.5\% each): office, multifamily, retail, industrial,
hotel, mixed-use, self-storage, medical office.
\item \textbf{Price range}: \$1M -- \$200M (discrete tiers: \$1--10M,
\$10--50M, \$50--200M, uniformly selected)
\item \textbf{Cap rate range}: 4\% -- 10\% (uniform)
\item \textbf{LTV range}: 55\% -- 80\% (target; actual LTV rounded to
nearest \$100K loan increment)
\item \textbf{Interest rate range}: 4.5\% -- 8.5\% (uniform, rounded to
4 decimal places)
\item \textbf{Amortization}: 25 or 30 years (uniformly selected)
\end{itemize}

Each scenario includes an input prompt phrased as a natural language
underwriting question, ground-truth computation markers with exact
input values and expected outputs, and complexity annotations (L1--L4).
All ground-truth values are computed using the same deterministic
functions used in the computation experts, ensuring zero systematic
bias in evaluation.

\subsection{Baseline Prompt Templates}
\label{sec:appendix:prompts}

\paragraph{Zero-Shot.}
\begin{quote}
\small
\textit{You are a commercial real estate underwriting analyst. Provide
precise calculations.}

[scenario prompt]
\end{quote}

\paragraph{Chain-of-Thought.}
\begin{quote}
\small
\textit{You are a commercial real estate underwriting analyst. Provide
precise calculations.}

\textit{Think step by step. Show all calculations clearly before giving
final answers.}

[scenario prompt]

\textit{Let me work through this step by step:}
\end{quote}

\paragraph{Structured Output.}
\begin{quote}
\small
\textit{You are a CRE underwriting analyst. Calculate ALL metrics from
the data below. Output ONLY the final numerical answers in this EXACT
format:}

\textit{DSCR: [number]}\\
\textit{LTV: [number]\%}\\
\textit{Cap Rate: [number]\%}\\
\textit{Debt Yield: [number]\%}

\textit{If a metric cannot be calculated, write N/A.}

[scenario prompt]

\textit{DSCR:}
\end{quote}

\subsection{Reproducibility}
\label{sec:appendix:reproduce}

All code, data generation scripts, training pipelines, and evaluation
scripts will be made publicly available upon publication. The complete
experiment (hidden state pre-computation, 5-fold probe training,
200-scenario evaluation with 3 baseline methods) completes in under
30 minutes on a single B200 GPU. Random seed: 42 for data splits;
seeds 0--2 for per-fold probe training.
