\section{Conclusion}
\label{sec:conclusion}

We introduced Hybrid MoE, a novel architecture that integrates
deterministic computation experts into the routing mechanism of sparse
Mixture-of-Experts transformers. By replacing neural FFN experts with a
three-stage pipeline (learned extraction, exact computation, learned
projection), we enable guaranteed-correct calculations within a single
forward pass.

Our experiments on gpt-oss-20b demonstrate that: (1)~financial values
are recoverable from transformer hidden states with 5-fold
cross-validated $R^2$ up to 0.98, (2)~routing signals achieve perfect
accuracy in detecting computation-requiring inputs, and (3)~the
approach yields a 173$\times$ speedup over single-sample generation.
With 5-fold CV ensembles, all four experts achieve 60--87\% accuracy
within 20\% error while guaranteeing a structured numerical output
for every input ($N{=}200$).

The key insight is that MoE routing provides a natural mechanism for
\emph{learned dispatch} between neural and symbolic computation: the
router learns when exact computation is needed, and the extraction MLP
learns how to read function arguments from dense representations. This
is fundamentally different from tool-calling approaches that require
explicit argument parsing and external execution.

Our analysis reveals two important findings. First, probe training
instability on small datasets, not hidden state quality, was
the primary cause of extraction failures; 5-fold CV ensembles resolve
this entirely. Second, error compounding through deterministic
computation (especially division) means that direct output probes can
outperform the extract-then-compute pipeline despite lower $R^2$,
suggesting that the hidden state encodes final answers at a useful
level of fidelity.

\paragraph{Future Work.}
Several directions are promising: (1)~attention-based extraction that
attends to specific number-bearing tokens rather than mean-pooling,
(2)~joint training of extraction and routing end-to-end through the
deterministic function, (3)~extension to chained computations where one
expert's output feeds another's input across layers, and
(4)~application to domains beyond finance where exact computation is
critical (pharmacokinetics, structural engineering, tax compliance).

The broader vision is a \emph{hybrid intelligence} paradigm where
transformers handle natural language understanding and reasoning while
deterministic modules handle exact computation, combined through the
same routing mechanism that already governs expert selection in modern
LLMs.
