\section{Method}
\label{sec:method}

\input{sections/figure_arch}

\subsection{Background: MoE Routing}
\label{sec:method:background}

In a standard MoE transformer layer, an input hidden state
$\mathbf{h} \in \mathbb{R}^d$ is processed by a router that selects
$k$ experts from a pool of $N$ experts. The router computes logits
$\mathbf{g} = \mathbf{W}_r \mathbf{h} + \mathbf{b}_r$ where
$\mathbf{W}_r \in \mathbb{R}^{N \times d}$, selects the top-$k$
indices, and normalizes the corresponding logit values into weights.
In gpt-oss-20b, the model uses $N{=}32$ experts, $k{=}4$ (top-4),
$d{=}2880$, and sigmoid normalization:
\begin{equation}
    w_i = \frac{\sigma(g_i)}{\sum_{j \in \text{top-}k} \sigma(g_j)},
    \quad i \in \text{top-}k(\mathbf{g})
\label{eq:routing}
\end{equation}
The output of the MoE layer is the weighted sum of expert outputs:
$\mathbf{o} = \sum_{i \in \text{top-}k} w_i \cdot E_i(\mathbf{h})$,
where each $E_i$ is a feed-forward network (typically SwiGLU).

\subsection{Hybrid MoE Architecture}
\label{sec:method:arch}

We extend the expert pool from $N$ to $N + M$ experts, where the
original $N{=}32$ neural experts remain frozen and $M{=}4$ new
\emph{deterministic computation experts} are added
(\Cref{fig:architecture}b). The router weight
matrix is extended from $\mathbb{R}^{N \times d}$ to
$\mathbb{R}^{(N+M) \times d}$:
\begin{equation}
    \mathbf{W}_r^{\text{ext}} =
    \begin{bmatrix}
        \mathbf{W}_r^{\text{orig}} & \text{(frozen)} \\
        \mathbf{W}_r^{\text{new}} & \text{(trainable)}
    \end{bmatrix},
    \quad
    \mathbf{b}_r^{\text{ext}} =
    \begin{bmatrix}
        \mathbf{b}_r^{\text{orig}} \\
        \mathbf{b}_r^{\text{new}}
    \end{bmatrix}
\label{eq:extended_router}
\end{equation}
The new router columns $\mathbf{W}_r^{\text{new}} \in \mathbb{R}^{M \times d}$
are initialized with small random values ($\sigma{=}0.01$), and the new
biases $\mathbf{b}_r^{\text{new}}$ are initialized to $-2.0$ to prevent
premature selection of computation experts before training.

\subsection{Deterministic Expert Module}
\label{sec:method:expert}

Each deterministic expert $E_c$ replaces the standard FFN with a
three-stage pipeline (\Cref{fig:architecture}c), maintaining the same
interface $\mathbb{R}^d \to \mathbb{R}^d$:

\paragraph{Stage 1: Parameter Extraction.}
A learned MLP $f_\theta: \mathbb{R}^d \to \mathbb{R}^p$ maps the
pooled hidden state to the $p$ numerical arguments required by the
target function:
\begin{equation}
    \hat{\mathbf{x}} = f_\theta(\bar{\mathbf{h}})
    = \text{MLP}_{\theta}(\bar{\mathbf{h}})
\label{eq:extraction}
\end{equation}
where $\bar{\mathbf{h}}$ is the mean-pooled hidden state across the
input sequence and $\hat{\mathbf{x}} \in \mathbb{R}^p$ contains the
extracted argument values (e.g., NOI and Annual Debt Service for the
DSCR expert). The extraction MLP consists of three linear layers with
SiLU activations and dropout (architecture:
$d \to 512 \to 128 \to p$).

\paragraph{Stage 2: Deterministic Computation.}
The extracted parameters are passed to an exact Python function
$\mathcal{F}: \mathbb{R}^p \to \mathbb{R}^q$ implemented in PyTorch:
\begin{equation}
    \mathbf{y} = \mathcal{F}(\hat{\mathbf{x}})
\label{eq:computation}
\end{equation}
Because $\mathcal{F}$ uses standard PyTorch operations (division,
multiplication, \texttt{torch.pow}), the computation graph is preserved
and gradients flow through the deterministic function via autograd.
For example, the DSCR expert computes
$\mathcal{F}(\hat{x}_1, \hat{x}_2) = \hat{x}_1 / \hat{x}_2$,
where $\hat{x}_1$ is the extracted NOI and $\hat{x}_2$ is the
extracted Annual Debt Service. When the extracted values are correct,
the output is \emph{exactly} correct; there is no neural
approximation.

\paragraph{Stage 3: Result Projection.}
A learned MLP $g_\phi: \mathbb{R}^q \to \mathbb{R}^d$ projects the
computation result back to the hidden dimension:
\begin{equation}
    \mathbf{o}_c = \alpha \cdot g_\phi(\mathbf{y})
\label{eq:projection}
\end{equation}
where $\alpha$ is a learnable scalar initialized to $0.01$ that
controls the initial contribution of the deterministic expert to the
residual stream. The projection MLP uses the architecture
$q \to 256 \to 512 \to d$.

\subsection{CRE Computation Experts}
\label{sec:method:cre}

We implement four deterministic experts for commercial real estate
underwriting, each computing a standard industry metric:

\begin{table}[h]
\centering
\small
\begin{tabular}{@{}llll@{}}
\toprule
\textbf{Expert} & \textbf{Formula} & \textbf{Inputs} & \textbf{Output} \\
\midrule
DSCR & $\text{NOI} / \text{ADS}$ & NOI, Annual Debt Service & Ratio \\
LTV & $(\text{Loan} / \text{Value}) \times 100$ & Loan amount, Property value & Percentage \\
Cap Rate & $(\text{NOI} / \text{Price}) \times 100$ & NOI, Purchase price & Percentage \\
Debt Yield & $(\text{NOI} / \text{Loan}) \times 100$ & NOI, Loan amount & Percentage \\
\bottomrule
\end{tabular}
\caption{Deterministic computation experts for CRE underwriting. Each
implements an exact financial formula as a differentiable PyTorch
function.}
\label{tab:experts}
\end{table}

\subsection{Training Procedure}
\label{sec:method:training}

Training proceeds in two phases with the base model \textbf{fully
frozen} throughout:

\paragraph{Phase 1: Pre-computation.}
We run all 3{,}000 training scenarios through the frozen gpt-oss-20b
model and cache both mean-pooled and last-token hidden states at layers
12, 16, and 20. This eliminates the need to run the 21B-parameter model
during probe training.

\paragraph{Phase 2: Probe training.}
Using the cached hidden states, we train:
(a)~\textbf{Extraction MLPs}, one per expert per input parameter,
trained to regress from the mean-pooled hidden state to the target
financial value in log-scale (using $\text{sign}(y) \cdot
\log(1 + |y|)$ to handle the wide dynamic range of financial values
from \$100K to \$200M). We use the Huber loss, AdamW optimizer
($\text{lr}{=}10^{-3}$, weight decay 0.01), and cosine annealing
over 300 epochs.
(b)~\textbf{Router signal classifier}, a binary classifier trained
to detect whether a given input requires deterministic computation,
trained with BCE loss for 30 epochs.

The total training involves ${\sim}5$M parameters and completes in
under 3 minutes on a single GPU.
