\section{Related Work}
\label{sec:related}

\paragraph{Sparse Mixture-of-Experts.}
MoE models route each token to a subset of experts, achieving
sub-linear inference cost relative to total parameters
\citep{shazeer2017outrageously,lepikhin2021gshard}. Switch Transformer
\citep{fedus2022switch} simplified routing to top-1 selection.
Mixtral \citep{jiang2024mixtral} demonstrated strong open-weight MoE
models with top-2 routing across 8 experts. DeepSeek-V2
\citep{deepseekai2024deepseekv2} introduced fine-grained experts and
shared expert isolation. Our base model, gpt-oss-20b, uses 32 experts
per layer with top-4 routing and sigmoid-normalized weights. To our knowledge, all prior MoE work uses exclusively neural or
trivial (zero/copy) experts; we introduce \emph{deterministic
computation} experts that execute exact domain functions.

\paragraph{MoE Expert Expansion.}
Several works have added experts to existing MoE models post-training.
Sparse Upcycling \citep{komatsuzaki2023sparse} converts dense models
to MoE by duplicating FFN layers as experts. Branch-Train-Merge
\citep{li2022branch} trains expert modules independently. MoE++
\citep{jin2024moepp} introduces heterogeneous expert types (zero
experts, copy experts) but these perform trivial operations. Nexus
\citep{gritsch2024nexus} adds task-specialized neural experts with
adaptive routing from domain embeddings. All of these add \emph{neural} experts.
Our deterministic experts are a fundamentally new class: they execute
exact symbolic computation rather than learned approximations.

\paragraph{Tool-Augmented Language Models.}
Toolformer \citep{schick2023toolformer} teaches models to insert API
calls into generated text via self-supervised learning. ReAct
\citep{yao2023react} interleaves reasoning and action traces for
multi-step tool use. ToolkenGPT \citep{hao2023toolkengpt} represents
tools as special vocabulary tokens, enabling tool invocation through
standard next-token prediction. Gorilla \citep{patil2023gorilla}
fine-tunes models on API documentation. These approaches share a common
limitation: tool execution is \emph{external} to the model, requiring
separate inference passes, API calls, or decoding steps. Hybrid MoE
embeds computation \emph{inside} the forward pass through the MoE
routing mechanism.

\paragraph{In-Model Arithmetic and Neurosymbolic Integration.}
The broader vision of combining neural and symbolic computation has a
long history \citep{garcez2019neural}. Two recent works are closely
related to ours. OccamLLM \citep{deleo2024occamllm} uses LLM hidden
states to control a symbolic architecture (OccamNet) that performs
exact arithmetic operations ($+, -, \times, \div, \sin$, etc.) in a
single autoregressive step, achieving 100\% accuracy on individual
operations. IGC \citep{dietz2025igc} integrates a gated calculator
directly into the LLM forward pass, achieving 98--99\% on the BigBench
Arithmetic benchmark without external tools. Both demonstrate that
hidden states contain sufficient information to drive exact computation.
Our work differs from these approaches in using \emph{MoE routing} for
dispatch: rather than adding a separate module or gating mechanism, we
extend the existing expert routing to naturally select between neural and
deterministic experts, which avoids introducing new architectural
components and allows the model to decide when computation is needed
through the same mechanism it already uses for expert selection.
Additional related work includes neural module networks
\citep{andreas2016neural} and dual-system neuro-symbolic approaches
\citep{nye2021improving,baydin2018automatic}.

\paragraph{Numerical Representations in LLMs.}
Understanding how transformers encode numbers is an active area.
\citet{wallace2019nlp} showed that MLPs can probe word embeddings for
numerical magnitudes. \citet{geva2023dissecting} demonstrated that
factual information is stored in specific feed-forward neurons.
\citet{chen2024states} found that LLMs develop ``implicit discrete
state representations'' for numerical reasoning, though these are ``far
from lossless.'' \citet{bai2025numbers} showed that LLMs encode number
values linearly in hidden states, recovering log-magnitudes with
${\sim}2.3\%$ error via linear probes. Most relevantly,
\citet{razeghi2022impact} showed that numerical reasoning accuracy
correlates with the frequency of specific numbers in pretraining data.
Our extraction experiments extend this line of work to domain-specific
financial values (NOI, loan amounts, property values), showing that
mean-pooled representations at middle layers encode these quantities
with $R^2$ up to 0.98.
